---
title: "Tmap"
format: html
---

We will map the HFMD cases notified in Pulau Pinang in 2022

# Packages installation and load

```{r}
library(tmap) ## thematic maps in R
library(sf) ## simple feature creation for mapping
library(readxl) ## to load data
library(here) ## read data
library(dplyr) ##data wrangling
```

# Part 1

## Importing the data

```{r}
data("World") #open source map in SF package
```

Now that we have the map, we can plot the "World" map by using function tm_shape. 
tm_shape will determine the layer of map that we wan to plot as base map. 

## tm_polygon determine which variable that we want to show on the chloropleth map

```{r}
tm_shape(World) +
  tm_polygons("life_exp")
```

World map produced, however the breaks and scales used were not really show any differences

## We need to modify it with (break = ) functions in the package.

```{r}
tm_shape(World) +
  tm_polygons("life_exp",
              breaks = c (0, 30, 40, 50, 60,70,80, Inf))
```

## Lets customize the map to make it more informative and reader's friendly

```{r}
tm_shape(World) +
  tm_polygons("life_exp",
              title = "Life Expectancy" , ## add title to the legend
              breaks = c (0, 30, 40, 50, 60,70,80, Inf)) +
    tm_borders() ## add borders on the map 

```

We can view it in dynamic map view
Just need to activate and shif the plotting mode into view mode

## Activate "view" mode

```{r}
tmap_mode("view")
```

Rerun the map
We can click on the map or zoom-in the area/ country
in this example, population density information will be appear once we click any country

## Plot the map
```{r}
tm_shape(World) +
  tm_polygons("life_exp",
              title = "Life Expectancy" , ## add title to the legend
              breaks = c (0, 30, 40, 50, 60,70,80, Inf)) +
    tm_borders() ## add borders on the map 

```

## we can edit the pop-up to make it more readers-friendly

```{r}
tm_shape(World) +
  tm_polygons("life_exp",
              title = "Life Expectancy" , ## add title to the legend
               breaks = c (0, 30, 40, 50, 60,70,80, Inf),
              id = "name",
              popup.vars = c("Country" = "name","Life_Expectancy" = "life_exp", "Pop_Density" = "pop_est_dens")) +
    tm_borders() ## add borders on the map 
```

# Part 2

## Lets map the disease

We need base map (area of interest) and data of the cases including the coordinates

## Load the case data

```{r}
Case <- read_excel("CASE.xlsx")
```

```{r}
glimpse(Case)
```

## Convert the case data to an SF object

```{r}
Case_1 <- st_as_sf(Case,coords = c("LONGITUDE","LATITUDE"), 
                    crs = 4326)  ## give coordinate referral system 
Case_1 <- st_transform(Case_1, 3168) ## Transform the CRS into Kertau format
```

## Plot the case

```{r}
tm_shape(Case_1) + 
  tm_dots(col = "black")
```

## Load the base map

```{r}
Penang <- st_read(here("Penang", 
                  "mukim2000_penang1.shp"))
```

## Give CRS to base map (similar to case data)

```{r}
st_crs(Penang)<-3168  ## Transform the CRS into Kertau format
```

## Plotting mode
```{r}
tmap_mode("plot") ## plotting mode
```

## Now that we have both cases and base map, we can plot the cases on the map

```{r}
tm_shape(Penang) +  ## The base map layer
  tm_polygons("NAMA_DP", 
              title = "HFMD Cases in Pulau Pinang 2022",
              title.size = 10) +
  tm_shape(Case_1) +  ## the case layer
  tm_dots(col = "black")+ ## use tm_dots to represent each case
  tm_layout(legend.outside = TRUE)
```

We can plot the cases according to specific district

## We just need to adjust the layer

```{r}
tm_shape(Penang[Penang$NAMA_DP == "SP.TENGAH",])+ ## specify the district
  tm_polygons("NAMMUK",
              title = "HFMD Cases in SP.TENGAH in 2022")+
  tm_shape(Case_1[Case_1$DAERAH == "SEBERG PERAI TENGAH",])+ ## specify the district in 2nd ayer
  tm_dots (col = "black") +  ## use tm_dots to represent each case
  tm_layout(legend.outside = TRUE)
```

We also can make direct comparison between groups (eg: gender) by using tm_facet function

## Tm_Facet function

```{r}
tm_shape(Penang) +  ## The base map layer
  tm_polygons("NAMA_DP", 
              title = "HFMD Cases in Pulau Pinang 2022",
              title.size = 10) +
  tm_shape(Case_1) +  ## the case layer
  tm_dots(col = "black")+   ## use tm_dots to represent each case
  tm_layout(legend.outside = TRUE) +
  tm_facets("GENDER", ncol = 2) 
```

Now, we can try it for areal data too (mapping the incidence rate)

## Lets load the data

```{r}
Case_2 <- read_excel( "Case_Sub.xlsx")
Case_3 <- merge(Penang, Case_2, by = c("NAMA_DP", "NAMMUK"))  
Incidence <- Case_3 %>% mutate(Incidence_Rate = (n/PENDUDUK)*1000)
```

```{r}
glimpse(Incidence)
```

## Plot the incidence

```{r}
tm_shape(Incidence) +
  tm_polygons("Incidence_Rate",
              title = "Incidence of HFMD Pulau Pinang 2022",
              title.size = 1,
             breaks = c (0,6,9,12,18, Inf)) +
    tm_compass(type = "4star", size = 2, position = c("left", "top")) + 
  tm_layout(legend.outside = TRUE)
```

## Lets make it dynamic

```{r}
tmap_mode("view")
```

```{r}
tm_shape(Incidence) +
  tm_polygons("Incidence_Rate",
              title = "Incidence of HFMD Pulau Pinang 2022",
              title.size = 1,
             breaks = c (0,6,9,12,18, Inf),
             id = "NAMA_DP",
              popup.vars = c("Subdistrict" = "NAMMUK","Case" = "n", "Incidence" = "Incidence_Rate")) +
    tm_borders() ## add borders on the map 
```
